---
title: "index.rmd"
author: "Yash Vig"
date: "2/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("httr")
library("jsonlite")
library("knitr")
library("dplyr")
```

## Summary Statistics Description
This gives a distribution of the programming experience of the corresponding to the number of students in each category

```{r echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
class.data <- read.csv(file = "data/intro-survey.csv", stringsAsFactors = FALSE)
programming.experience <- data.frame(table(class.data$Programming.Experience))
colnames(programming.experience) <- c("Programming Experience", "Number of People")
kable(programming.experience)
```

This gives a similar distribution to the table above, but instead it shows the distribtuion of experience when it comes to command line programming.

```{r echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
commandline.experience <- data.frame(table(class.data$Command.Line.Experience))
colnames(commandline.experience) <- c("Command-Line Experience", "Number of People")
kable(commandline.experience)
```

Here we see that thereis no direct corelations between the distributions of the two types of programming.